---
title: "CDatanet: An  R package to simulate and estimate a Count Data Model with Social Interactions"
author: "Elysée Aristide Houndetoungan"
date: "`r Sys.Date()`"
output:
  pdf_document:
    citation_package: natbib
    number_sections: true
  bookdown::pdf_book:
    citation_package: biblatex
bibliography: ["References.bib", "Packages.bib"]
biblio-style: "apalike"
link-citations: true
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{CDatanet Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

There is a large and growing literature on peer effects in economics.^[For recent reviews, see @boucher2016some, @de2017econometrics and @bramoulle2019peer.] Recent contributions include among other, models for discrete data such as binary [e.g., @lee2014binary; @liu2019simultaneous], ordered [e.g., @boucher2018peer] and multinomial [e.g., @guerra2017multinomial]. However, there are hardly any models dealing with count data (e.g., data collected by asking respondents: "*How many times do you smoke daily?*"). 


I present a model in which the outcome takes count values (0, 1, 2, \dots). In this model, individuals interact through a directed network and their peers' decision may influence their own decision. Moreover, they do not observe their peers' outcome and form rational belief about that.  

I also present an easy-to-use R package---named  **CDatanet**---for implementing the model. This note is a supplement for the package manual. I provide examples with simulated data to show how to use the package. The package is located on GitHub at [github.com/ahoundetoungan/CDatanet](https://github.com/ahoundetoungan/CDatanet). All the results of the note can be replicated following the code provided.

The rest of this note is organized as follow. Section [2](#model) briefly introduces the model. Section [3](#package) explains how to install the package. Sections [4](#exo) and [5](#endo) provide examples in which the network matrix is assumed exogenous and endogenous respectively. They present Monte Carlo experiments to assess the performance of the estimator of the model parameters. They also discuss how to control for the endogeneity of the network. Section [6](#con) concludes.

# Model{#model}
Let $\mathcal{V} = \{1, \dots, n\}$ be a set of $n$ players indexed by $i$ and $y_i$ the observed integer outcome of player $i$ (e.g., the number of cigarettes smoked per day or per week). I denote by $\boldsymbol{G}$ the adjacency matrix such that $G_{ij} = \dfrac{1}{n_i}$ if $i$ knows $j$ and $G_{ij} = 0$ otherwise, where $n_i$ the number of friends of $i$. Let also $\boldsymbol{X}$ be the matrix of explanatory variables. 

As in @liu2019simultaneous, I assume that there is a latent variable $y_i^*$ which determines the observed count variable $y_i$ . This latent variable $y_i^*$ and the observed $y_i$ are linked as follow.

Let be $\left(a_q\right)_{q \in \mathbb{N}}$ a sequence  given by $a_0 = -\infty$, $a_1 \in \mathbb{R}$ and $a_q = a_1 + \gamma(q-1)$ for $q \in \mathbb{N}^*$ and $\gamma \in \mathbb{R}_+^*$. If $y_i^* \in (a_{q}, a_{q + 1}]$ then $y_i = q$.

The first order conditions of the game equilibrium implies that

$$ y_i^*  =  \lambda\sum_{i = 1}^n \sum_{r = 0}^{\infty} rG_{ij} p_{ir}  + \boldsymbol{x}_i^{\prime}\boldsymbol{\beta} + \varepsilon_{i}$$
where $p_{ir}$ is the probability of $y_i = r$, $\lambda$ is the peer effect and $\varepsilon_{i} \sim \mathcal{N}\left(0, \sigma_{\varepsilon}^2\right)$.

The game equilibrium is given by
$$ p_{ir} = \Phi\left(\dfrac{\lambda\sum_{i = 1}^n \sum_{r = 0}^{\infty} rG_{ij} p_{ir}  + \boldsymbol{x}_i^{\prime}\boldsymbol{\beta} - a_q}{\sigma_{\varepsilon}}\right) - \Phi\left(\dfrac{\lambda\sum_{i = 1}^n \sum_{r = 0}^{\infty} rG_{ij} p_{ir}  + \boldsymbol{x}_i^{\prime}\boldsymbol{\beta}  - a_{q+1}}{\sigma_{\varepsilon}}\right)$$
where $\Phi$ is the probability density function of $\mathcal{N}(0, 1)$.

To estimate $\boldsymbol{\theta} = (\lambda, \boldsymbol{\beta}^{\prime}, \sigma_{\varepsilon})^{\prime}$, I rely on the NPL method. I refer the reader to @houndetoungan2020cdata for more details. The estimation method is implemented in the package **CDatanet**.

# **CDatanet** Installation {#package}
The code source of the package is written in C++ with the package **Rcpp** [@R-Rcpp]. **CDatanet**  is currently available on GitHub. Its installation requires to prior install **devtools** package [@R-devtools]. Moreover, Windows users should install Rtools compatible with their R version.

**CDatanet** package can be installed from this GitHub repos using the `install_github` function of **devtools**. All the dependencies will also be installed automatically.

```{r install, echo = TRUE, eval = FALSE}
library(devtools)
install_github("ahoundetoungan/CDatanet")
```

# Examples with exogenous network{#exo}
In this section, I present examples to show how to use the package. I simulate data following the model by assuming that the network matrix is exogenous. I then show how to estimate the model parameters using functions implemented in **CDatanet**. I also use Monte Carlo simulations to assess the performance of the estimator of the model parameters.

## Data simulation{#exo.data}
Given the adjacency matrix, the explanatory variables and $\boldsymbol{\theta}$, the function `simCDnet` can used to simulate data. I assume that there are `M` sub-networks. The number of individuals in each sub-network is randomly chosen between 100 and 1000. I assume two exogenous variables (plus the intercept) and I control for the contextual effect.

```{r dataexo, echo = TRUE, eval = TRUE}
set.seed(2020) 
library(CDatanet)
# Groups' size
M      <- 5 # Number of sub-groups
nvec   <- round(runif(M, 100, 1000))
print(nvec)
n      <- sum(nvec)
print(n)

# Parameters
lambda <- 0.4              # peer effect
beta   <- c(2, -1.9, 0.8)  # own effects
gamma  <- c(1.5, -1.2)     # contextual effects
sigma  <- 1.5              # standard deviation of epsilon
theta  <- c(lambda, beta, gamma, sigma)
  
# X
data   <- data.frame(x1 = rnorm(n, 1, 1), x2 =  rexp(n, 0.4))
```

To simulate the network matrix, I assume that the number of friends of each individual is randomly chosen between 0 and 30.

```{r netexo, echo = TRUE, eval = TRUE}
# Network
Glist  <- list()
for (m in 1:M) {
  nm           <- nvec[m]
  Gm           <- matrix(0, nm, nm)
  max_d        <- 30
  for (i in 1:nm) {
    tmp        <- sample((1:nm)[-i], sample(0:max_d, 1))
    Gm[i, tmp] <- 1
  }
  rs           <- rowSums(Gm); rs[rs == 0] <- 1
  Gm           <- Gm/rs
  Glist[[m]]   <- Gm
}
```

I can now simulate the count data. The output of `simCDnet` includes `yst` the vector of $y_i^*$, `y` the vector of $y_i$, `yb` the vector of $\displaystyle\sum_{r = 0}^{\infty}rp_{ir}$, `Gyb` the vector of $\displaystyle\sum_{i = 1}^n \sum_{r = 0}^{\infty} rG_{ij} p_{ir}$ and `iteration` as the number of iterations performed to find the fixed point `yb`.

```{r dataexo2, echo = TRUE, eval = TRUE, fig.height = 3, fig.align = "center"}
ytmp    <- simCDnet(formula = ~ x1 + x2 | x1 + x2, Glist = Glist, theta = theta,
                    data = data)
names(ytmp)
y       <- ytmp$y
# Add y to data
data$y  <- y
# Summarize y
summary(y)
table(y)
# Plot data histogram
library(ggplot2)
print(ggplot(data = data.frame(y = y), aes(x = y)) +
        geom_bar(color = "black", fill = "cyan") +
        theme_bw() + xlab("") + ylab("Frequency"))
```
## Estimation
Using the simulated data, I estimate the count data model as well as the SART and SAR models. The output of these models has a `summary class` to summarize the results. In addition, the `summary` function computes marginal effects for the count data model.
```{r estexo, echo = TRUE, eval = TRUE}
#  Count data
CD   <- CDnetNPL(formula = y ~ x1 + x2, contextual = TRUE, Glist = Glist,
                  optimizer = "nlm", 
                 data = data, npl.ctr = list(print = FALSE, maxit = 5e3))
summary(CD)

# SART
SART   <- SARTML(formula = y ~ x1 + x2, contextual = TRUE, Glist = Glist,
                optimizer = "nlm", data = data, print = FALSE)
summary(SART)

#SAR
SAR  <- SARML(formula = y ~ x1 + x2, contextual = TRUE, Glist = Glist,
              optimizer = "nlm", data = data, print = FALSE)
summary(SAR)
```

Using the count data model, the peer effect is estimated at 0.38, which is close to the true value, 0.4. By replicating this experience several times, the average of the estimates is likely to be almost equal to the true value (see next section). Note that this parameter is not interpretable. Instead, one should interpret the marginal effect: *increasing of the expected $y_j$ of the peers by one implies an increasing of the expected $y_i$ of the individual by 0.28*. However, estimating the SART model using data simulated from the count data model underestimates the peer effect (0.27). The package does not implement the marginal effect under the SART model. This latter should be much lower. Lastly, the SAR model more underestimates the peer effect (0.16). In addition, this is directly the marginal effect because the SAR model does not assume a limited dependent variable.

## Monte Carlo Simulations
In this section, I conduct Monte Carlo simulations to assess the performance of the estimator. I use the same settings as in Section [4.1](#exo.data) except that the number of individuals in each sub-network is set to 500. Let's build the following Monte Carlo function that simulates the data, estimates the three models and returns their estimators.

```{r mcfun, echo = TRUE, eval = TRUE}
fMC <- function(s) {
  # Groups' size
  M      <- 5
  nvec   <- rep(500, 5); 
  n      <- sum(nvec)
  # Parameters
  lambda <- 0.4 
  beta   <- c(2, -1.9, 0.8)
  gamma  <- c(1.5, -1.2)
  sigma  <- 1.5   
  theta  <- c(lambda, beta, gamma, sigma)
  
  # X
  data   <- data.frame(x1 = rnorm(n, 1, 1), x2 =  rexp(n, 0.4))
  
  # Network
  Glist  <- list()
  for (m in 1:M) {
    nm           <- nvec[m]
    Gm           <- matrix(0, nm, nm)
    max_d        <- 30
    for (i in 1:nm) {
      tmp        <- sample((1:nm)[-i], sample(0:max_d, 1))
      Gm[i, tmp] <- 1
    }
    rs           <- rowSums(Gm); rs[rs == 0] <- 1
    Gm           <- Gm/rs
    Glist[[m]]   <- Gm
  }
  
  # y
  ytmp   <- simCDnet(formula = ~ x1 + x2 | x1 + x2, Glist = Glist, theta = theta,
                     data = data)
  y      <- ytmp$y
  data$y <- y
  
  #  Models
  CD     <- CDnetNPL(formula = y ~ x1 + x2, contextual = TRUE, Glist = Glist,
                     optimizer = "nlm", 
                     data = data, npl.ctr = list(print = FALSE, maxit = 5e3))
  SART   <- SARTML(formula = y ~ x1 + x2, contextual = TRUE, Glist = Glist,
                   optimizer = "nlm", data = data, print = FALSE)
  SAR    <- SARML(formula = y ~ x1 + x2, contextual = TRUE, Glist = Glist,
                  optimizer = "nlm", data = data, print = FALSE)
  c(CD$estimate, SART$estimate, SAR$estimate)
}
```

I run the Monte Carlo function `fMC` 1000 times and compute the average of each estimator. To make it faster, I run the replications in parallel using **doParallel** package [@R-doParallel]. 

```{r mc, echo = TRUE, eval = FALSE, message = FALSE}
library(doParallel)
n.cores  <- 32
replic   <- 1000 #Number of replications
out.mc   <- mclapply(1:replic, fMC, mc.cores = n.cores)
out.mc   <- apply(t(do.call(cbind, out.mc)), 2, mean)
out.mc   <- cbind(theta, out.mc[1:7], out.mc[8:14], out.mc[15:21])
colnames(out.mc) <- c("TrueValue", "CoutData", "SART", "SAR")
print(out.mc)
```
```{r mcprint, echo = FALSE, eval = TRUE}
load("out.mc.Rdata")
print(out.mc)
```

The estimator of the count data model seems unbiased while the SART and the SAR models underestimate the peer effect.

# Examples with endogenous network{#endo}
Peer effect estimation is generally based on the assumption of exogeneity of the adjacency matrix. This means that the probability of link formation is not correlated to the error term in the count data model. Such an assumption is strong as the link formation probability may depend on unobserved characteristics (e.g, gregariousness) that also influence the outcome. In this section, I simulate data with endogenous network (which violates the assumption). I then show that the peer effect is overestimated when one does not control for the endogeneity of the network. I also present a method to control for the endogeneity. Finally, I use Monte Carlo simulations to prove that this method performs well.

## Data simulation
I assume three sub-networks. The number of individuals in each sub-network is randomly chosen between 100 and 500.
```{r dataendo, echo = TRUE, eval = TRUE}
rm(list = ls())
set.seed(2020) 
# Groups' size
M      <- 3 # Number of sub-groups
nvec   <- round(runif(M, 100, 500))
print(nvec)
n      <- sum(nvec)
print(n)

# Parameters
lambda <- 0.4               # peer effect
beta   <- c(2, -1.9, 0.8)   # own effects
gamma  <- c(1.5, -1.2)      # contextual effects
sigma  <- 1.5               # standard deviation of epsilon
theta  <- c(lambda, beta, gamma, sigma)

# X
data   <- data.frame(x1 = rnorm(n, 1, 1), x2 =  rexp(n, 0.4))
```
The network matrix follows the dyadic linking model presented in @houndetoungan2020cdata. Let $\boldsymbol{A}$ be the matrix of links such that $A_{ij} = 1$ if $i$ knows $j$ and $A_{ij} = 0$. In other words, $A_{ij} = 1$ if $G_{ij} = \dfrac{1}{n_i}$. It is easier to work with $\boldsymbol{A}$ (binary data) in the network formation model. However, in the count data model, I use $\boldsymbol{G}$ as row-normalized equivalent of $\boldsymbol{A}$.

The probability of link formation, $P_{ij}$ depends on $\Delta x_{1ij} = |x_{1i} - x_{1j}|$ and $\Delta x_{2ij} = |x_{2i} - x_{2j}|$ (observed dyad-specific variables) as well as on unobserved individual-level attributes $\mu_i$ and $\mu_j$. Typically, let $a_{ij}^*$ defined by
$$a_{ij}^* = \Delta \boldsymbol{x}_{ij}^{\prime}\bar{\boldsymbol{\beta}} + \mu_i + \mu_j + \varepsilon_{ij}^*,$$
where $\varepsilon_{ij}^* \sim logistic$ distribution. I assume that $A_{ij} = 1$ if $a_{ij}^* > 0$. In that case, $P_{ij}$ is given by
$$P_{ij} = \dfrac{\exp{\left(\Delta \boldsymbol{x}_{ij}^{\prime}\bar{\boldsymbol{\beta}} + \mu_i + \mu_j\right)}}{1 + \exp{\left(\Delta \boldsymbol{x}_{ij}^{\prime}\bar{\boldsymbol{\beta}} + \mu_i + \mu_j\right)}}.$$
I also assume that $\mu_i$ is random. If $i$ is from the s-th sub-network, then $\mu_i \sim \mathcal{N}\left(u_{\mu s}, \sigma_{\mu s}^2\right)$. The mean and the variance of $\mu_i$ vary across the sub-networks. This is a way to control for the sub-network heterogeneity as fixed effects in the link formation probability.
```{r netendo, echo = TRUE, eval = TRUE}
# Parameter for network model
betanet  <- c(-2.8, -1.5)    # beta
Glist    <- list()           # adjacency matrix row normalized
Network  <- list()           # adjacency matrix row non-normalized
dX       <- matrix(0, 0, 2)  # observed dyad-specific variables
mu       <- list()           # unobserved individual-level attribute
uu       <- runif(M, -1, 1)  # mean of uu in each sub-network
sigma2u  <- runif(M, 0.5, 4) # variance of uu in each sub-network

# Network
for (m in 1:M) {
  nm           <- nvec[m]
  mum          <- rnorm(nm, uu[m], sqrt(sigma2u[m]))
  Z1           <- matrix(0, nm, nm)  
  Z2           <- matrix(0, nm, nm)
  
  for (i in 1:nm) {
    for (j in 1:nm) {
      Z1[i, j] <- abs(data$x1[i] - data$x1[j])
      Z2[i, j] <- abs(data$x2[i] - data$x2[j])
    }
  }
  
  Gm           <- 1*((Z1*betanet[1] + Z2*betanet[2] +
                        kronecker(mum, t(mum), "+") + rlogis(nm^2)) > 0)
  diag(Gm)     <- 0
  
  diag(Z1)     <- NA
  diag(Z2)     <- NA
  Z1           <- Z1[!is.na(Z1)]
  Z2           <- Z2[!is.na(Z2)]
  
  dX           <- rbind(dX, cbind(Z1, Z2))
  
  Network[[m]] <- Gm
  rs           <- rowSums(Gm); rs[rs == 0] <- 1
  Gm           <- Gm/rs
  Glist[[m]]   <- Gm
  mu[[m]]      <- mum
}
mu            <- unlist(mu)
```

To simulate the dependent variable, I use $\boldsymbol{\mu}$, the vector of $\mu_i$ as additional explanatory variable in the count variable model. I set that $Cor(\mu_i, \varepsilon_i) = \rho$ and $Cor(\sum_{j = 1}^nG_{ij}\mu_i, \varepsilon_i) = \bar{\rho}$. In that case, $\varepsilon_{i} = \rho \sigma_{\varepsilon}\tilde{\mu}_i + \bar{\rho} \sigma_{\varepsilon}\bar{\tilde{\mu}}_i  + \tilde{\nu_i}$, where $\tilde{\mu}_i = \dfrac{\mu_i - u_{\mu s}}{\sigma_{\mu s}}$, $\bar{\tilde{\mu}}_i = \sum_{j = 1}^nG_{ij}\tilde{\mu}_j$ is the average of $\tilde{\mu}_i$ among $i$'s friends and $\tilde{\nu_i}$ is the new error term following normal distribution of variance $\bar{\sigma}^2_{\varepsilon} = \left(1 - \rho^2 - \bar{\rho}^2\right)\sigma^2_{\varepsilon}$.
```{r dataendo2, echo = TRUE, eval = TRUE, fig.height = 3, fig.align = "center"}
tmu      <- (mu - rep(uu, nvec))/sqrt(rep(sigma2u, nvec))
data$tmu <- tmu
rho      <- 0.24
rhobar   <- 0.18
thetanet <- c(lambda, beta, sigma*rho, gamma, sigma*rhobar,
              sigma*(1 - rho - rhobar))
ytmp     <- simCDnet(formula = ~ x1 + x2 + tmu | x1 + x2 + tmu,
                     Glist = Glist, theta = thetanet, data = data)
y        <- ytmp$y
# Add y to data
data$y   <- y
# Summarize y
summary(y)
# Plot data histogram
print(ggplot(data = data.frame(y = y), aes(x = y)) +
        geom_bar(color = "black", fill = "cyan") +
        theme_bw() + xlab("") + ylab("Frequency"))
```

In real life, $\tilde{\mu}_i$ and $\bar{\tilde{\mu}}_i$ are not observed and they are included in the error term $\varepsilon_i$. This implies that the network is endogenous.

## Estimation
When the endogeneity of the network is not taken into account, the peer effect is overestimated. The reason is that the peer effect also captures other effects. The model considers any common shock on $\tilde{\mu}_i$ and $\bar{\tilde{\mu}}_i$ as peer effect. In the example below, the peer effect is estimated at 0.49 when I do not control for the endogeneity of the network.
```{r estendo1, echo = TRUE, eval = TRUE}
# Count data model
CDexo <- CDnetNPL(formula = y ~ x1 + x2, contextual = TRUE, Glist = Glist,
                  optimizer = "optim", data = data, npl.ctr = list(print = FALSE))
summary(CDexo)
```

To correct the endogeneity, $\tilde{\mu}_i$ and $\bar{\tilde{\mu}}_i$ should be included in the model as additional explanatory variables. However, these variables are not observed.

I use Markov Chain Monte Carlo (MCMC) to estimate the parameters of the dyadic linking model (these include the unobserved individual-level attributes $\mu_i$). The estimation can be done using the function `netformation`. 
```{r estendo2, echo = TRUE, eval = F}
# Dyadic linking model
net <- netformation(network =  Network, formula = ~ dX, fixed.effects = TRUE,
                    mcmc.ctr = list(burnin = 1000, iteration = 5000))
```

I randomly choose some parameters and present their posterior distribution. It stands out that the MCMC converges quickly and the simulations are pretty close to the true values.

```{r estendo2a, echo = TRUE, eval = F}
# plot simulations
par(mfrow = c(4,2), mar = c(2, 2, 2, 1.9))
plot(net$posterior$beta[,1], type = "l", ylim = c(-2.9, -2.6), col = "blue",
     main = bquote(beta[1]), ylab = "")
abline(h = betanet[1], col = "red")

plot(net$posterior$beta[,2], type = "l", ylim = c(-1.6, -1.4), col = "blue",
     main = bquote(beta[2]), ylab = "")
abline(h = betanet[2], col = "red")

plot(net$posterior$mu[,10], type = "l", col = "blue",
     main = bquote(mu[10]), ylab = "")
abline(h = mu[10], col = "red")

plot(net$posterior$mu[,542], type = "l", col = "blue",
     main = bquote(mu[542]), ylab = "")
abline(h = mu[542], col = "red")

plot(net$posterior$mu[,10], type = "l", col = "blue",
     main = bquote(mu[10]), ylab = "")
abline(h = mu[10], col = "red")

plot(net$posterior$mu[,849], type = "l", col = "blue",
     main = bquote(mu[849]), ylab = "")
abline(h = mu[849], col = "red")

plot(net$posterior$mu[,752], type = "l", col = "blue",
     main = bquote(mu[752]), ylab = "")
abline(h = mu[752], col = "red")

plot(net$posterior$sigmamu2[,3], type = "l", col = "blue",
     main =  bquote(sigma[mu][3]^2), ylab = "")
abline(h = sigma2u[3], col = "red")
```

Using simulations from the posterior distribution, I can construct good estimators for $\tilde{\mu}_i$ and $\bar{\tilde{\mu}}_i$ and use them as additional supplementary explanatory variables. In the next example, I use the simulation with the highest posterior density to estimate $\tilde{\mu}_i$ and $\bar{\tilde{\mu}}_i$.
```{r estendo3, echo = TRUE, eval = F}
t           <- which.max(net$posterior$log.density)
print(t)
muest       <- net$posterior$mu[t,]
uuest       <- net$posterior$uu[t,]
sigma2uest  <- net$posterior$sigmamu2[t,]
tmuest      <- (muest - rep(uuest, nvec))/sqrt(rep(sigma2uest, nvec))
data$tmuest <- tmuest
CDendo      <- CDnetNPL(formula = y ~ x1 + x2 + tmuest, contextual = TRUE, 
                        Glist = Glist, optimizer = "optim", data = data, 
                        npl.ctr = list(print = FALSE))
summary(CDendo)
```
The coefficients of $\tilde{\mu}_i$ and $\bar{\tilde{\mu}}_i$ are significant. This confirms that the network is endogenous. Moreover, the peer effect is estimated at 0.40, which is much better than the previous estimate. 

The standard errors computed in this example are not valid because they assume that $\tilde{\mu}_i$ and $\bar{\tilde{\mu}}_i$ are observed. To correct the standard error, I replicate simulations of $\tilde{\mu}_i$ and $\bar{\tilde{\mu}}_i$ with replacement from the posterior distribution in order to take into account the variance of the MCMC. 

```{r estendo4, echo = TRUE, eval = FALSE}
fendo    <- function(s) {
  t         <- sample(3001:8000, 1)
  datat     <- data
  mus       <- net$posterior$mu[t,]
  uus       <- rep(net$posterior$uu[t,], nvec)
  sus       <- rep(net$posterior$sigmamu2[t,], nvec)
  tmu       <- (mus - uus)/sqrt(sus)
  datat$tmu <- tmu
  
  CDnet    <- CDnetNPL(formula = y ~ x1 + x2 + tmu, contextual = TRUE, 
                       Glist = Glist, optimizer = "optim", data = datat,
                       npl.ctr = list(print = FALSE))
  
  summary(CDnet, Glist = Glist, data = datat)
}

n.cores    <- 32
replic     <- 1000 #Number of replications
out.endo   <- mclapply(1:replic, fendo, mc.cores = n.cores)
# The output of out.endo is a list of objects from "summary.CDnetNPL" class
# Let's set the class of out.endo as "summary.CDnetNPLs"
class(out.endo) <- "summary.CDnetNPLs"
# I can now summarize the results using print
# Object from "summary.CDnetNPL" has a print method
print(out.endo)
```

```{r endoprint1, echo = FALSE, eval = TRUE}
load("out.endo.Rdata")
CDatanet:::print.summary.CDnetNPLs(out.endo)
```
The standard errors are slightly greater than the previous estimate. 

## Monte Carlo Simulations
One simulation is not sufficient to confirm that the method used to correct the endogeneity performs well. In this section, I use Monte Carlo simulations to assess the performance of this method. I compare the estimates when the network is assumed exogenous to these when I control for the endogeneity.

I build the following Monte Carlo function which simulates data based on endogenous network, and estimates the model by assuming that the network is exogenous and then endogenous.
```{r mcfunendo, echo = TRUE, eval = TRUE}
fMCendo <- function(s) {
  # parameters
  M        <- 3 
  nvec     <- rep(500, 3)
  n        <- sum(nvec)
  
  lambda   <- 0.4              # peer effect
  beta     <- c(2, -1.9, 0.8)  # own effects
  gamma    <- c(1.5, -1.2)     # contextual effects
  sigma    <- 1.5              # standard deviation of epsilon
  theta    <- c(lambda, beta, gamma, sigma)
  
  data     <- data.frame(x1 = rnorm(n, 1, 1), x2 =  rexp(n, 0.4))
  
  # Network
  betanet  <- c(-2.8, -1.5)    # beta
  Glist    <- list()           # adjacency matrix row normalized
  Network  <- list()           # adjacency matrix row non-normalized
  dX       <- matrix(0, 0, 2)  # observed dyad-specific variables
  mu       <- list()           # unobserved individual-level attribute
  uu       <- runif(M, -1, 1)  # mean of uu in each sub-network
  sigma2u  <- runif(M, 0.5, 4) # variance of uu in each sub-network
  
  for (m in 1:M) {
    nm           <- nvec[m]
    mum          <- rnorm(nm, uu[m], sqrt(sigma2u[m]))
    Z1           <- matrix(0, nm, nm)  
    Z2           <- matrix(0, nm, nm)
    
    for (i in 1:nm) {
      for (j in 1:nm) {
        Z1[i, j] <- abs(data$x1[i] - data$x1[j])
        Z2[i, j] <- abs(data$x2[i] - data$x2[j])
      }
    }
    
    Gm           <- 1*((Z1*betanet[1] + Z2*betanet[2] +
                          kronecker(mum, t(mum), "+") + rlogis(nm^2)) > 0)
    diag(Gm)     <- 0
    
    diag(Z1)     <- NA
    diag(Z2)     <- NA
    Z1           <- Z1[!is.na(Z1)]
    Z2           <- Z2[!is.na(Z2)]
    
    dX           <- rbind(dX, cbind(Z1, Z2))
    
    Network[[m]] <- Gm
    rs           <- rowSums(Gm); rs[rs == 0] <- 1
    Gm           <- Gm/rs
    Glist[[m]]   <- Gm
    mu[[m]]      <- mum
  }
  mu             <- unlist(mu)
  
  # Data
  tmu      <- (mu - rep(uu, nvec))/sqrt(rep(sigma2u, nvec))
  data$tmu <- tmu
  rho      <- 0.24
  rhobar   <- 0.18
  thetanet <- c(lambda, beta, sigma*rho, gamma, sigma*rhobar,
                sigma*(1 - rho - rhobar))
  ytmp     <- simCDnet(formula = ~ x1 + x2 + tmu | x1 + x2 + tmu,
                       Glist = Glist, theta = thetanet, data = data)
  y        <- ytmp$y
  data$y   <- y
  
  # Count data model
  CDexo <- CDnetNPL(formula = y ~ x1 + x2, contextual = TRUE, Glist = Glist,
                    optimizer = "optim", data = data, npl.ctr = list(print = FALSE))
  # Dyadic linking model
  net   <- netformation(network =  Network, formula = ~ dX, fixed.effects = TRUE,
                        mcmc.ctr = list(burnin = 1000, iteration = 2000), print = FALSE)
  # Endogeneity
  t           <- which.max(net$posterior$log.density)
  muest       <- net$posterior$mu[t,]
  uuest       <- net$posterior$uu[t,]
  sigma2uest  <- net$posterior$sigmamu2[t,]
  tmuest      <- (muest - rep(uuest, nvec))/sqrt(rep(sigma2uest, nvec))
  data$tmuest <- tmuest
  CDendo      <- CDnetNPL(formula = y ~ x1 + x2 + tmuest, contextual = TRUE, 
                          Glist = Glist, optimizer = "optim", data = data, 
                          npl.ctr = list(print = FALSE))
  c(CDexo$estimate, CDendo$estimate)
}
```

I now run the function `fMCendo` 1000 times and compute the average of both estimators. 

```{r mcendoout, echo = TRUE, eval = FALSE, message = FALSE}
n.cores              <- 5
replic               <- 1000 #Number of replications
out.mcendo           <- mclapply(1:replic, fMCendo, mc.cores = n.cores)
out.mcendo           <- apply(t(do.call(cbind, out.mcendo)), 2, mean)
out.mcendo           <- cbind(thetanet, c(out.mcendo[c(1:4, NA, 5:6, NA, 7)]), 
                              out.mcendo[8:16])
colnames(out.mcendo) <- c("TrueValue", "EndoNotControlled", "EndoControlled")
print(out.mcendo)
```
```{r mcprintendo, echo = FALSE, eval = FALSE}
load("out.mcendo.Rdata")
print(out.mcendo)
```

The results confirm that the method used to correct the endogeneity performs well. The average of the estimates is -- whereas the true value is --. In contrast, the peer effect is overestimated when the network is assumed exogenous. The average of the estimates is --. 

# Conclusion {#con}
This paper provides technical details on the package **CDatanet**. It shows with simple and practical examples how to use the package. It provides comparison of the model with the SART and the SAR models. It also shows with Monte Carlo simulations how the model performs when the network matrix is exogenous and endogenous. Simulation results prove that the peer effect is overestimated when the network is endogenous.  Lastly, it presents a way to control for the endogeneity of the network. 